---
title: 'Reproducible report: <br/> Phenotypic and molecular characterization of the resistance to strobilurins in Fusarium graminearum'
output:
  html_document:
    css: style.css
    depth: 5
    number_sections: no
    theme: default
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
date: 'Created: 2021-10-06 <br/>Last update: `r Sys.Date()`'
---

<link href="https://fonts.googleapis.com/css?family=Oswald" rel="stylesheet">



### R Packages

All packages loaded during this work.

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(ec50estimator)
library(cowplot)
library(ggthemes)
library(drc)
library(ggridges)
library(epiR)
library(car)
library(gsheet)
library(patchwork)
library(rnaturalearth)
library(ggmap)
library(ggrepel)
library(lme4)
library(lmerTest)
library(ggsn)

```

### Isolates 

All isolates used in this study with their location were loaded to plotting the map below.

```{r}
data_load = gsheet2tbl("https://docs.google.com/spreadsheets/d/1YTmns-Mjyz6HWluKN5ksHCpJfD_jshavGVI52oLfszU/edit#gid=1862629108") 

data_load
```

### By year and state

```{r}
map_all  = data_load %>% 
  group_by(safra, estado) %>%
  summarise(n = length(Isolate)) 
map_all
```

### By year in RS

```{r}

p1 <- map_all %>% 
  filter(estado=="RS") %>% 
ggplot(aes(factor(safra), y= n))+
  ylim(0,60)+
  geom_col( position = "dodge",width = 0.5, color="black",fill="black", alpha=0.6)+
  scale_fill_colorblind()+
  theme_minimal()+
  labs(x = "Year", y = "Count of Isolates", fill="")

ggsave("img/n_year_rs.png",width = 5, height = 3, dpi = 300)
```

### By year in PR


```{r}

p2 <- map_all %>% 
  filter(estado=="PR") %>% 
ggplot(aes(factor(safra), y= n))+
  geom_col( position = "dodge",width = 0.5, color="black",fill="orange", alpha=0.6)+
  scale_fill_colorblind()+
  theme_minimal()+
  labs(x = "Year", y = "Count of Isolates", fill="")

ggsave("img/n_year_pr.png",width = 5, height = 3, dpi = 300)
```



```{r}
#install.packages("devtools") 
#devtools::install_github("ropensci/rnaturalearthhires")
library("rnaturalearth")

BRA <- ne_states(country = "Brazil",
  returnclass = "sf")

```

### By municipality 

```{r}
map_all_1  = data_load %>% 
  group_by(estado, município) %>%
  summarise(n = length(Isolate),
            lat=mean(lat), 
            long= mean(long)) 
map_all_1
```

###  Map of isolates

```{r}

p3 <- ggplot() +
  geom_sf(data = BRA, fill = "white",color = "black")+
geom_point(data = map_all_1, aes(as.numeric(long),as.numeric(lat), color= estado, fill= estado, size=n), shape = 21, width = 0.9)+
geom_text_repel(data = map_all_1, aes(as.numeric(long),as.numeric(lat), label=(n)), size = 4, box.padding = 0.2, min.segment.length =1, seed = F, max.overlaps = Inf)+
  xlim(-58,-48)+
  ylim(-34,-23)+
  #scale_size_continuous(breaks = seq(1,10, by = 1), range = c(1, 6))+
  #facet_wrap(~safra,nrow = 2)+
  theme_map()+
  scale_color_manual(values = c("orange","black"))+
  scale_fill_manual(values = c("orange","black"))+
  guides(color=F, size=F)+
  
 # theme(legend.position = "none")+
  labs(fill="", size="", x="", y="", color="" )+
   annotate("text", x=-53, y=-24, label="PR", size=4)+
  annotate("text", x=-54, y=-30, label="RS", size=4)
 p3
  ggsave("img/map_new_5.png", dpi= 600, height = 8, width = 8 )
```

```{r}
library(patchwork)
(p3 | (p2/p1)) +
  plot_annotation(tag_levels = "A")+
  plot_layout(widths = c(2, 1))
ggsave("img/figure1.png", width =8, height =5)
  
```



In this study, we phenotypically characterize Fusarium graminearum isolates to azoxystrobin and pyraclostrobin. 



### EC50 - Pyra

## Data

This database contains information about all germination tests conducted, divided by: experiment and geographic collection.

```{r}
germination_assay<- gsheet2tbl("https://docs.google.com/spreadsheets/d/13uMSrA-gLXXhkgTKxrE0MjZX1ewLZGxnTUIpWCD-n64/edit#gid=1516403512") %>% 
  group_by(Isolate, Dose, Replicates, Trail) %>%
  dplyr::mutate(GC = as.numeric(GC)) %>%
  dplyr::mutate(Dose = as.numeric(Dose))
  
germination_assay
```

## Visualization


```{r}

germination_assay %>% 
  ggplot(aes(as.factor(Dose), GC, color=as.factor(Trail)))+
    stat_summary(fun.data = "mean_cl_boot",
               geom = "errorbar",
               width = 0,
               size =.5,
               position = position_dodge(width=.5))+
  stat_summary(fun = "mean",
               size = 1.4,
               geom = "point",
               position = position_dodge(width=.5))+
  facet_wrap(~Isolate)+
  scale_color_colorblind()+
  theme_minimal_hgrid(font_size = 6)+
   labs(x = expression(Dose~~μg/ml), 
        y = "Number of germinated conidia/dose", color="Trial")

```


## Models  

Ten isolates were chosen to select the best model that fits all the data set.   

Isolates: "09Tr006"

```{r}
g1<-germination_assay %>% 
  filter(Isolate=="07Tr037")
g1
  
```


```{r}
model_1=drm(GC~Dose, data=g1, fct = W1.3())
plot(model_1)
summary(model_1)
ED(model_1,50)
```




```{r}
mselect(model_1,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"UEM-2665"

```{r}
g2<-germination_assay %>% 
  filter(Isolate=="08Tr005")
g2
```



```{r}
model_2=drm((GC)~Dose, data=g2, fct = W1.3())
plot(model_2)
summary(model_2)
ED(model_2,50)
```


```{r}
mselect(model_2,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```
"09Tr006"

```{r}
g3<-germination_assay %>% 
  filter(Isolate=="09Tr006")
g3
```

```{r}
model_3=drm((GC)~Dose, data=g3, fct = LL.4())
plot(model_3)
summary(model_3)
ED(model_3,50)
```


```{r}
mselect(model_3,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"10Tr042"

```{r}
g4<-germination_assay %>% 
  filter(Isolate=="10Tr042")
g4
```


```{r}
model_4=drm((GC)~Dose, data=g4, fct = W1.3())
plot(model_4)
summary(model_4)
ED(model_4,50)
```


```{r}
mselect(model_4,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"11Tr005"

```{r}
g5<-germination_assay %>% 
  filter(Isolate=="11Tr005")
g5
```


```{r}
model_5=drm((GC)~Dose, data=g5, fct = W1.3())
plot(model_5)
summary(model_5)
ED(model_5,50)
```

```{r}
mselect(model_5,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"3072"

```{r}
g6<-germination_assay %>% 
  filter(Isolate=="3072")
g6
```


```{r}
model_6=drm((GC)~Dose, data=g6, fct = W1.3())
plot(model_6)
summary(model_6)
ED(model_6,50)
```




```{r}
mselect(model_6,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"3708"

```{r}
g7<-germination_assay %>% 
  filter(Isolate=="3708")
g7
```

```{r}
model_7=drm((GC)~Dose, data=g7, fct = W1.3())
plot(model_7)
summary(model_7)
ED(model_7,50)
```


```{r}
mselect(model_7,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"3712"

```{r}
g8<-germination_assay %>% 
  filter(Isolate=="3712")
g8
```


```{r}
model_8=drm((GC)~Dose, data=g8, fct = W1.3())
plot(model_8)
summary(model_8)
ED(model_8,50)
```

```{r}
mselect(model_8,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"UEM-2628"

```{r}
g9<-germination_assay %>% 
  filter(Isolate=="UEM-2628")
g9
```


```{r}
model_9=drm((GC)~Dose, data=g9, fct = LL.3())
plot(model_9)
summary(model_9)
ED(model_9,50)
```


```{r}
mselect(model_9,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"UEM-3687"

```{r}
g10<-germination_assay %>% 
  filter(Isolate=="UEM-3687")
g10
```





```{r}
model_10=drm((GC)~Dose, data=g10, fct = W1.3())
plot(model_10)
summary(model_10)
ED(model_10,50)
```


```{r}
mselect(model_10,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(),BC.4(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```



## EC50 Pyra

Calculation of Ec50 for the number of germinated conidia in relation to the dose, being stratified by year and population (RS, PR). The model was chosen through the tests performed above considering <IAC and best curve fit.

Model: W1.3


```{r}
ec50_ger=estimate_EC50(GC~Dose,
                data= germination_assay ,
                isolate_col= "Isolate",
                strata_col = c("Year", "Population"),
                fct=drc::W1.3(),
                interval="delta")
ec50_ger

```

### Statistics 

PARANÁ - PR

```{r}
ec50_ger %>% 
  #filter(Population=="PR") %>% 
  summarise(mean=mean(Estimate),
            median=median(Estimate),
            sd=sd(Estimate),
            lower= quantile(Estimate, 0.25),
            upper= quantile(Estimate, 0.75))
   
  
  
```

RIO GRANDE DO SUL- RS

```{r}
ec50_ger %>% 
  filter(Population=="RS") %>% 
  summarise(mean=mean(Estimate),
            median=median(Estimate),
            sd=sd(Estimate),
            lower= quantile(Estimate, 0.25),
            upper= quantile(Estimate, 0.75))
   
  
  
```


ALL ISOLATES 

```{r}
ec50_ger %>% 
  filter(Population=="PR") %>% 
  summarise(mean=mean(Estimate),
            median=median(Estimate),
            sd=sd(Estimate),
            lower= quantile(Estimate, 0.25),
            upper= quantile(Estimate, 0.75))
   
  
  
```


##Anova


```{r}
aov_Ger<- lm(sqrt(Estimate)~Population,data= ec50_ger)
summary(aov_Ger)
Anova(aov_Ger)
```

## Homocedasticity

```{r}
leveneTest(Estimate~Population,data=ec50_ger)
```

##Normality

```{r}
shapiro.test(residuals(aov_Ger))
```




## Boxplot Ec50 

```{r}
box_pyra = ec50_total_pyra=ec50_ger %>% 
  filter(Estimate<100) %>% 
  ggplot(aes(Population, log(Estimate)))+
  ylim(-6, 6)+
  geom_jitter(size=2, color = "black", width = 0.05)+
  geom_boxplot(width=0.5, outlier.color = NA, color= "black", fill = NA, size =1)+
  scale_color_colorblind()+
  #geom_hline(yintercept = .50, color="red")+
  #annotate("text", x=2.5,y=0.55, label="DD", size=4, color="red")+
  #facet_wrap("Population")+
  theme_minimal_hgrid()+
   labs(y = "log of EC50 ", 
        x="")
ggsave("img/box_PYRA.png", dpi = 500, height = 3.5, width = 5)

box_pyra
```


### EC50 Azox 

## Data

This database contains information about all germination tests conducted, divided by: experiment and population.

```{r}
germination_assay_azox<-  gsheet2tbl("https://docs.google.com/spreadsheets/d/1Mis_iDuVrq8x_IpjJDdhAeNPKHiIAZfea9sljgI1sIw/edit#gid=0") %>% 
  group_by(Isolate, Dose, Replicates, Trial) %>%
  dplyr::mutate(GC = as.numeric(GC)) %>%
  dplyr::mutate(Dose = as.numeric(Dose))
  
germination_assay_azox
```

##Visualization


```{r}

germination_assay_azox %>% 
  ggplot(aes(as.factor(Dose), GC, color=as.factor(Trial)))+
    stat_summary(fun.data = "mean_cl_boot",
               geom = "errorbar",
               width = 0,
               size =.5,
               position = position_dodge(width=.5))+
  stat_summary(fun = "mean",
               size = 1.4,
               geom = "point",
               position = position_dodge(width=.5))+
  facet_wrap(~Isolate)+
  scale_color_colorblind()+
  theme_minimal_hgrid(font_size = 6)+
   labs(x = expression(Dose~~μg/ml), y = "Number of germinated conidia/dose", color="")

```




##Models  

Seven isolates were chosen to select the best model that fits all the data set.   

Isolates: "07Tr012"



```{r}
a1<-germination_assay_azox %>% 
  filter(Isolate=="07Tr012")
a1
  
```



```{r}
az1=drm(GC~Dose, data=a1, fct = W2.3())
plot(az1)
summary(az1)
ED(az1, 50)
```



```{r}
mselect(az1,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"09Tr019"

```{r}
a2<-germination_assay_azox %>% 
  filter(Isolate=="09Tr019")
a2
  
```



```{r}
az2=drm(GC~Dose, data=a2, fct = W1.3())
plot(az2)
ggsave("img/az2.png", dpi = 500)
summary(az2)
ED(az2, 50)


```



```{r}
mselect(az2,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"11Tr074"

```{r}
a3<-germination_assay_azox %>% 
  filter(Isolate=="11Tr074")
a3
  
```


```{r}
az3=drm(GC~Dose, data=a3, fct = W1.3())
plot(az3)
summary(az3)
ED(az3, 50)
```



```{r}
mselect(az3,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"3070"

```{r}
a4<-germination_assay_azox %>% 
  filter(Isolate=="3070")
a4
  
```


```{r}
az4=drm(GC~Dose, data=a4, fct = W1.3())
plot(az4)
summary(az4)
ED(az4, 50)
```

```{r}
mselect(az4,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"UEM-3688"

```{r}
a5<-germination_assay_azox %>% 
  filter(Isolate=="UEM-3688")
a5
  
```




```{r}
az5=drm(GC~Dose, data=a5, fct = W2.3())
plot(az5)
summary(az5)
ED(az5, 50)
```




```{r}
mselect(az5,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"UEM-2580"

```{r}
a6<-germination_assay_azox %>% 
  filter(Isolate=="UEM-2580")
a6
  
```


```{r}
az6=drm(GC~Dose, data=a6, fct = W2.3())
plot(az6)
summary(az6)
ED(az6, 50)
```


```{r}
mselect(az6,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

"3711"

```{r}
a7<-germination_assay_azox %>% 
  filter(Isolate=="3711")
a7
  
```


```{r}
az7=drm(GC~Dose, data=a7, fct = W2.3())
plot(az7)
summary(az7)
ED(az7, 50)
```


```{r}
mselect(az7,fctList =list(LL.2(),LL.3(), LL.3u(),LL.4(),LL.5(),W1.2(), W1.3(),W1.3u(), W1.4(),W2.4(), W2.3(), W2.2(), BC.5(),LL2.2(), LL2.3(), LL2.3u(),LL2.4(),LL2.5(), AR.2(), AR.3(), MM.2(), MM.3()))
```

## EC50 Azox

Calculation of Ec50 for the number of germinated conidia in relation to the dose, being stratified by year and population (RS, PR). The model was chosen through the tests performed above considering <IAC and best curve fit.

Model: W1.3



```{r}
ec50_ger_AZX=estimate_EC50(GC~Dose,
                data= germination_assay_azox ,
                isolate_col= "Isolate",
                strata_col = c("Year", "Population"),
                fct=drc::W1.3(),
                interval="delta")
ec50_ger_AZX

```


### Statistics

PARANÁ - PR

```{r}
ec50_ger_AZX %>% 
  filter(Population=="PR") %>% 
  summarise(mean=mean(Estimate),
            median=median(Estimate),
            sd=sd(Estimate),
            lower= quantile(Estimate, 0.25),
            upper= quantile(Estimate, 0.75))
   
  
  
```



RIO GRANDE DO SUL - RS 


```{r}
ec50_ger_AZX %>% 
  filter(Population=="RS") %>% 
  summarise(mean=mean(Estimate),
            median=median(Estimate),
            sd=sd(Estimate),
            lower= quantile(Estimate, 0.25),
            upper= quantile(Estimate, 0.75))
   
  
  
```


ALL ISOLATES

```{r}
ec50_ger_AZX %>% 
  summarise(mean=mean(Estimate),
            median=median(Estimate),
            sd=sd(Estimate), 
            lower_azox= quantile(Estimate, 0.25),
            upper_azox= quantile(Estimate, 0.75))
   
  
  
```



##Anova


```{r}
aov_Ger_azox<- lm((Estimate)~Population,data= ec50_ger_AZX)
summary(aov_Ger_azox)
Anova(aov_Ger_azox)
```

## Homocedasticity

```{r}
leveneTest(Estimate~Population,data=ec50_ger_AZX)
```

##Normality

```{r}
shapiro.test(residuals(aov_Ger_azox))
```

## GLM  


```{r}
glm_azox = glm(Estimate ~ Population , family = Gamma (link=log), data  = ec50_ger_AZX )
summary(glm_azox)

```





##Boxplot Ec50 


```{r}
box_azox_2=ec50_total_azox=ec50_ger_AZX %>% 
  ggplot(aes(Population, log(Estimate)))+
  geom_jitter(size=2, color = "black", width = 0.05)+
  ylim(-6,6)+
  geom_boxplot(width=0.5, outlier.color = NA, color= "black", fill = NA, size =1)+
  scale_color_colorblind()+
   #geom_hline(yintercept = 5, color="red")+
  #annotate("text", x=2.5,y=5.5, label="DD", size=4, color="red")+
  #facet_wrap("Population")+
  theme_minimal_hgrid()+
   labs(y = "Log of EC50",
        x="")
ggsave("img/box_AZOX_2.png", dpi = 500, height = 3.5, width = 5)

box_azox_2
```







## EC50  (both fungicides)

Here we loaded in the same data frame the EC50 values for both fungicides. 

```{r}
ec50_fun= gsheet2tbl("https://docs.google.com/spreadsheets/d/116vNPmNBOEoH0oWIFZxV9B8gkkv-wcYJ_fhbW2rpBZU/edit#gid=1135537345")

ec50_fun
```

## Boxplot (azox and pyra)

```{r}

box_azox_pyra <- ec50_fun %>% 
  filter(Estimate<100) %>% 
  ggplot(aes(Fungicide, log(Estimate)))+
  geom_boxplot(width=0.5, outlier.color = NA,  size =1)+
  ylim(-6, 6)+
  geom_jitter(size=2, color = "black",fill="black", width = 0.05)+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank())+
  theme_minimal_hgrid()+
  scale_fill_few()+
   labs(x = "Fungicides",
        y= "Log of EC50")+
   theme(legend.position = "none")




```



## Figure 2


```{r fig.height=4, fig.width=8}
(box_azox_2 | box_pyra)+
  plot_annotation(tag_levels = 'A')
ggsave("img/figure2.png", width = 8, height =5)

```

## Correlation EC50

```{r}

cor_ec50= gsheet2tbl("https://docs.google.com/spreadsheets/d/116vNPmNBOEoH0oWIFZxV9B8gkkv-wcYJ_fhbW2rpBZU/edit#gid=0") %>% 
  mutate(pyra=log(pyra)) %>% 
  mutate(azox=log(azox))
cor_ec50
```

```{r}

cor.test((cor_ec50$azox), (cor_ec50$pyra), method ="spearman")
```



```{r}
cor_ec50 %>% 
  ggplot(aes(pyra, azox))+
  #geom_smooth(method=lm)+
  geom_point(shape=21, size=3)+
  theme_minimal()+
  labs(x =expression((Log)~EC[50]~~μg/ml~Pyraclostrobin), y = expression((Log)~EC[50]~~μg/ml~Azoxystrobin, color=""))

ggsave("img/cor_ec50_log.png", dpi = 500)
```






### ISOLATE CML3066  

We also calculated the EC50 for the isolates CML3066, which is a isolate from Brazil that has its genome sequenced. 

```{r}

EC50_P<- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Mis_iDuVrq8x_IpjJDdhAeNPKHiIAZfea9sljgI1sIw/edit#gid=1606966266") %>% 
   filter(Code =="GEN-P") %>% 
   dplyr::mutate(GC = as.numeric(GC)) %>%
  dplyr::mutate(Dose = as.numeric(Dose)) 
 
  
EC50_P
```




## EC50 CML3066 Pyra

```{r}
model_p=drm(GC~Dose, data=  EC50_P, fct = W1.3())
plot(model_p)
summary(model_p)
ED(model_p,50)
```




### EC50- CML3066 Azox

```{r}

EC50_AZ<- gsheet2tbl("https://docs.google.com/spreadsheets/d/1Mis_iDuVrq8x_IpjJDdhAeNPKHiIAZfea9sljgI1sIw/edit#gid=1606966266") %>% 
   filter(Code =="GEN") %>% 
   dplyr::mutate(GC = as.numeric(GC)) %>%
  dplyr::mutate(Dose = as.numeric(Dose)) 
 
  
EC50_AZ
```




## EC50 CML3066 Azox

```{r}
model_az=drm(GC~Dose, data=  EC50_AZ, fct = W1.3())
plot(model_az)
summary(model_az)
ED(model_az,50)
```


### Discriminatory Dose

We also screening isolates through discriminatory dose for each fungicide. 

## Load data RS


```{r}
data_load_rs = gsheet2tbl("https://docs.google.com/spreadsheets/d/1xaHgZCiO85EwuowMlGKlvO1tlwG2SleaBG-L-5x4Enw/edit#gid=0") %>% 
  mutate(Fungicide = factor(Fungicide, levels = c("Control", "Azoxtrobin","Pyraclostrobin" )))
data_load_rs
```

## visualization 

```{r}
data_load_rs %>% 
  ggplot(aes(Fungicide,conidia))+
  geom_jitter()
```
## Single isolate

```{r}
single_rs  = data_load_rs %>% 
  filter(Isolate == 1)  
single_rs %>% 
  ggplot(aes(Fungicide,conidia, color = factor(Trial)))+
  geom_jitter(height = 0)
  
```

### Glm

```{r}
glm_single = glm(conidia ~ Fungicide , family = poisson, data  = single_rs)
summary(glm_single)

coef(glm_single)[2:3]
confint(glm_single)

azox_single = c(exp(coef(glm_single)[2]), exp(confint(glm_single)[2,]))
pyra_single = c(exp(coef(glm_single)[3]), exp(confint(glm_single)[3,]))
```



### Data frame (GLM results)

```{r}
isolates = unique(data_load_rs$Isolate)
azox_all = NULL
pyra_all = NULL

for(i in 1:length(isolates)){
single_df  = data_load_rs %>% 
  filter(Isolate == isolates[i]) 

glm_single = glm(conidia ~ Fungicide , family = poisson, data  = single_df)
summary(glm_single)

coef(glm_single)[2:3]
confint(glm_single)

azox_single = c(mean = as.numeric(exp(coef(glm_single)[2])), exp(confint(glm_single)[2,]), isol = isolates[i], year = unique(single_df$Year), local = unique(single_df$City),lat = unique(single_df$lat), long = unique(single_df$long), state= (single_df$State))
pyra_single = c(mean = as.numeric(exp(coef(glm_single)[3])), exp(confint(glm_single)[3,]), isol = isolates[i], year = unique(single_df$Year),local = unique(single_df$City), lat = unique(single_df$lat), long = unique(single_df$long), state= (single_df$State))

azox_all = rbind(azox_all, azox_single)
pyra_all = rbind(pyra_all, pyra_single)

}

results_rs = bind_rows(
  
as.data.frame(azox_all) %>% 
  mutate(fungicide = "Azoxystrobin"),

as.data.frame(pyra_all) %>% 
  mutate(fungicide = "Pyraclostrobin")

)
```


## Data frame reorganization 

```{r}
results_rs_1<-results_rs %>%  
summarise(mean=as.numeric(mean), 
         lower=as.numeric(`2.5 %`),
         upper= as.numeric(`97.5 %`),
        lat=as.numeric(lat),
        long=as.numeric(long), 
        isolate= as.numeric(isol),
        fungicide=fungicide, 
        year=year, 
        state=state1) 
results_rs_1
```


## Boxplot (all isolates)
```{r}
box_1=results_rs_1 %>% 
  ggplot(aes(fungicide, mean), color=lower>0.5)+
  ylim(0,1)+
  geom_jitter(size=2, width = 0.2, color="gray")+
  geom_boxplot(width=0.5, outlier.color = NA, color= "black", fill = NA, size =1)+
  scale_color_colorblind()+
   geom_hline(yintercept = 0.5, color="red")+
  #annotate("text", x=2.5,y=5.5, label="DD", size=4, color="red")+
  #facet_wrap("Population")+
  theme_minimal()+
   labs(y = "",
        x="")

box_1
#ggsave("img/box_AZOX.png", dpi = 500, height = 3.5, width = 5)


```


## Less sensitive year, Pyra, RS

```{r fig.height=5, fig.width=5}
results_rs_2 <- results_rs_1 %>% 
  filter(fungicide== "Pyraclostrobin")

dd_1=results_rs_2 %>% 
  ggplot(aes(year, mean, group=lower, color=lower>0.5))+
  geom_hline(yintercept = .50, color="red")+
  scale_color_colorblind()+
  scale_y_continuous(limits = c(0, 1.2))+
  geom_jitter(position= position_dodge(width = 1))+
  geom_errorbar(aes(ymin = lower, ymax = upper, color = lower>0.5), width=1, position= position_dodge(width = 1), size=0.3)+
  #facet_wrap(~fungicide, ncol = 1)+
theme_minimal(base_size = 18)+
   labs(x="", y="", color="LL IC95")+
  theme(legend.position = "none")


  
dd_1
ggsave("img/dd_rs_2.png",dpi= 600, height = 8, width = 8)
```

## Less sensitive, year, Azox, RS

```{r fig.height=5, fig.width=5}
results_rs_3 <- results_rs_1 %>% 
  filter(fungicide== "Azoxystrobin")

dd_2=results_rs_3 %>% 
  ggplot(aes(year, mean, group=lower, color=lower>0.5))+
  geom_hline(yintercept = .50, color="red")+
  scale_color_colorblind()+
  scale_y_continuous(limits = c(0, 1.2))+
  geom_jitter(position= position_dodge(width = 1))+
  geom_errorbar(aes(ymin = lower, ymax = upper, color = lower>0.5), width=1, position= position_dodge(width = 1), size=0.3)+
  theme_minimal(base_size = 18)+
  labs(x="Year", y="", color="LL IC95")+
  theme(legend.position = "none")

  
dd_2

```




## Proportion by year 

```{r}
results_freq<-results_rs_1 %>% 
group_by( fungicide, year) %>% 
  mutate(n_isolates=length(isolate)) %>% 
  filter(lower>0.5) %>% 
  summarise(count=n(), 
            prop=(count/mean(n_isolates))*100)
  


results_freq


```

## Less sensitive, year, RS

```{r fig.height=10, fig.width=12}
dd_2+dd_1+plot_annotation(tag_levels = list(c("A", "B")))+plot_layout(ncol = 2)+
  labs(y= "")

ggsave("img/dd_all_all.png",dpi= 500, height = 8, width = 14)
```




## Load data - Paraná


```{r}
data_load_pr = gsheet2tbl("https://docs.google.com/spreadsheets/d/1xaHgZCiO85EwuowMlGKlvO1tlwG2SleaBG-L-5x4Enw/edit#gid=29082389") %>% 
  mutate(Fungicide = factor(Fungicide, levels = c("Control", "Azoxtrobin","Pyraclostrobin" ))) %>% 
  filter(!is.na(conidia))
data_load_pr
```



## visualization

```{r}
data_load_pr %>% 
  ggplot(aes(Fungicide,conidia))+
  geom_jitter()
```

## Single isolate

```{r}
single_pr  = data_load_pr %>% 
  filter(Isolate == 6040)  
single_pr %>% 
  ggplot(aes(Fungicide,conidia, color = factor(Trial)))+
  geom_jitter(height = 0)
  
```

### Glm

```{r}
glm_single_pr = glm(conidia ~ Fungicide , family = poisson, data  = single_pr)
summary(glm_single_pr)

coef(glm_single_pr)[2:3]
confint(glm_single_pr)

azox_single_pr = c(exp(coef(glm_single_pr)[2]), exp(confint(glm_single_pr)[2,]))
pyra_single_pr = c(exp(coef(glm_single_pr)[3]), exp(confint(glm_single_pr)[3,]))
```

## Data frame Paraná

```{r}
isolates = unique(data_load_pr$Isolate)
azox_all_pr = NULL
pyra_all_pr = NULL

for(i in 1:length(isolates)){
single_df_pr  = data_load_pr %>% 
  filter(Isolate == isolates[i]) 

glm_single_pr = glm(conidia ~ Fungicide , family = poisson, data  = single_df_pr)
summary(glm_single_pr)

# coef(glm_single_pr)[2:3]
# confint(glm_single_pr)

azox_single_pr = c(mean = as.numeric(exp(coef(glm_single_pr)[2])), exp(confint(glm_single_pr)[2,]), isol = isolates[i], year = unique(single_df_pr$Year), state= (single_df_pr$State))
pyra_single_pr = c(mean = as.numeric(exp(coef(glm_single_pr)[3])), exp(confint(glm_single_pr)[3,]), isol = isolates[i], year = unique(single_df_pr$Year), state= (single_df_pr$State))

azox_all_pr = rbind(azox_all_pr, azox_single_pr)
pyra_all_pr = rbind(pyra_all_pr, pyra_single_pr)

}

results_pr = bind_rows(
  
as.data.frame(azox_all_pr) %>% 
  mutate(fungicide = "Azoxystrobin"),

as.data.frame(pyra_all_pr) %>% 
  mutate(fungicide = "Pyraclostrobin")

)
results_pr
```


## Data frame reorganization

```{r}
results_pr_1<-results_pr %>%  
summarise(mean=as.numeric(mean), 
         lower=as.numeric(`2.5 %`),
         upper= as.numeric(`97.5 %`),
        isolate= as.numeric(isol),
        fungicide=fungicide, 
        year=year, 
        state= state1) 
head(results_pr_1)
```

## Boxplot (all isolates)

```{r}
box_2=results_pr_1 %>% 
  ggplot(aes(fungicide, mean), color=lower>0.5)+
  ylim(0,1)+
  geom_jitter(size=2, width = 0.2, color="gray")+
  geom_boxplot(width=0.5, outlier.color = NA, color= "black", fill = NA, size =1)+
  scale_color_colorblind()+
   geom_hline(yintercept = 0.5, color="red")+
  #annotate("text", x=2.5,y=0.6, label="DD", size=4, color="red")+
  #facet_wrap("Population")+
  theme_minimal()+
   labs(y = "",
        x="Fungicides")
box_2
#ggsave("img/box_AZOX.png", dpi = 500, height = 3.5, width = 5)


```


```{r fig.height=10, fig.width=12}
box_1+box_2+plot_annotation(tag_levels = list(c("A", "B")))+plot_layout(ncol = 2)+
  labs(y= "")

ggsave("img/box_all.png",dpi= 500, height = 8, width = 12)
```

## Join data frame

Here we loaded in the same data frame all values for the discriminatory dose for both fungicides.

```{r}
DD_ALL= full_join(results_rs_1, results_pr_1) 

head(DD_ALL)
```

## Boxplot DD

```{r}
DD_ALL %>% 
  ggplot(aes(state, mean, fill= fungicide))+
  ylim(0,1)+
  geom_boxplot(width=0.5, outlier.color = NA, size =1,  alpha=0.3)+
    geom_jitter(size=1, width = 0.2, color="black")+
  scale_fill_manual(values = c("15","gray90"))+
  #scale_color_colorblind()+
   geom_hline(yintercept = 0.5, color="red")+
  #annotate("text", x=2.5,y=0.6, label="DD", size=4, color="red")+
  facet_wrap(~fungicide)+
  theme_minimal()+
   theme(legend.position = "none")+
   labs(y = "Prop of germinated conidia on the DD in relation to control ",
        x="Collection")
ggsave("img/box_all.png", dpi = 500, height = 6, width = 6)


```







