---
title: "Ejercicios 1"
---


1 - Mancha en red en cebada. 

3 variedades
10 hojas con sus valores de severidad

Calcule la incidencia y severidad media de cada trt 

```{r}
cebada <- rio::import("data/ceb.csv")
```

## Canola phoma

Experimento de canola conducido en Balcarce, donde fueron testeados 10 fungicidas (mas un control sin protección con fungicida) con 3 bloques en que se registró el progreso de manchas foliares de Phoma lingam a través del tiempo (tiempo térmico desde la d

- calcule el AUC de cada parcela

```{r}
canola <- readxl::read_excel("data/canola_maculas.xlsx")

canola %>%  
  pivot_longer(inc_15:inc_248, 
         names_to = "tt", 
         values_to = "inc", 
         names_prefix = "inc_")-> can_long

can_long %>%
  group_by(trt, bk) %>%
  summarize(AUC = MESS::auc(inc, tt))
```

## Downy mildew on grapes. 

The experiment was conducted at Wooster, Ohio, on the experimental farm of the Ohio Agricultural Research and Development Center, Ohio State University.

There were 3 blocks with 6 treatments. Treatment 1 is the unsprayed control. On 30 Sep 1990, disease incidence was measured. For each plot, 5 randomly chosen shoots on each of the 3 vines were observed. The canopy was closed and shoots could be intertwined. On each shoot, the total number of leaves and the number of infected leaves were recorded.

Used with permission of Larry Madden.

- Compare los trt (desconsidere vine y shoot) mediante GLMM

```{r}
grapes <- agridat::hughes.grapes
```
